---
title: "R Notebook"
output: html_notebook
---


```{r}
df <- read.csv('Energy_Usage_From_DOE_Buildings.csv') #Reading in data
```


Exploring data

```{r}
head(df)
```

Removing unneeded columns for my analysis

```{r}
df$X. <- NULL
df$X_ <- NULL
```

```{r}
str(df)
```

Removing more unneeded columns

```{r}
df$May.12 <- NULL
df$Jun.12 <- NULL
df$BIN <- NULL
df$BBL <- NULL
df$Community.Board <- NULL
df$Council.District <- NULL
df$Census.Tract <- NULL
```

1- Manhattan,  2- Bronx,  3- Brooklyn, 4 - Queens, 5- Staten Island, 6 - Citywide, 7- Out of City

What the difference borough numbers means.
```{r}
df$Borough[df$Borough == 1] <- 'Manhattan'
df$Borough[df$Borough == 2] <- 'Bronx'
df$Borough[df$Borough == 3] <- 'Brooklyn'
df$Borough[df$Borough == 4] <- 'Queens'
df$Borough[df$Borough == 5] <- 'Staten Island'
df$Borough[df$Borough == 6] <- 'Citywide'
df$Borough[df$Borough == 7] <- 'Out of City'
```


Renaming column to a better name.
```{r}
df$area_name <- df$NTA
df$NTA <- NULL
```


Removing all rows that doesnt measure energy usage
```{r}
df <- df[df$Measurement == 'Electricity Usage (KWH)',]
```


Checking how many unique building adresses there are
```{r}
length(unique(df$Building.Address))
```

Creating all the dates for my new dataframe
```{r}
dates <- c(rep(seq(from = as.Date("2008-07-01"), to = as.Date("2012-04-01"), by = 'month'), 36))
```


Getting all the adressnames in order so it fits with the dates for my new dataframe
```{r}
l <- c()
l <- as.list(l)
i=1
i <- 1
while(i <= length(unique(df$Building.Address))){
  b <- rep(unique(df$Building.Address)[[i]], 46)
  l <- c(l,b)
  i=i+1
}
```


Creating the new dataframe with all the dates
```{r}
new_df <- as.data.frame(dates, axis=2)
```

Adding all the adresses to the new dataframe
```{r}
new_df$adress <- l
```

Adding all the columns i want in the new df
```{r}
new_df$electricity_costs_usd <- NA
new_df$borough <- NA
new_df$area_name <- NA
new_df$postcode <- NA
new_df$latitude <- NA
new_df$longitude <- NA
```



Getting all the electricity costs per month from the old dataframe to the new dataframe

```{r}
l <- c()
l <- as.list(l)
i <- 1

while(i <= length(unique(df$Building.Address))){
  j <- 1
  while(j <= 46){ #There are only 46 months of data
    b <- t(df[i,4:49])[[j]]
    l <- c(l,b)
    j=j+1
  }
  i=i+1
}

new_df$electricity_costs_usd <- l #Adding the data to my new dataframe
```

Now it's time to add all the other missing values

```{r}
df$adress <- df$Building.Address #Making the both dataframes have the same name for adress
df$Building.Address <- NULL #Removing the old column
```

```{r}
tmp <- merge(df, new_df, by = "adress")

new_df$area_name <- tmp$area_name.x
new_df$latitude <- tmp$Latitude
new_df$longitude <- tmp$Longitude
new_df$postcode <- tmp$Postcode
new_df$borough <- tmp$Borough
```

Removing unwanted data and renaiming new_df to df

```{r}
df <- new_df
```

```{r}
remove(new_df)
remove(l)
remove(tmp)
remove(b)
remove(dates)
remove(i)
remove(j)
```

Converting df columns to the right types
```{r}
df$adress <- as.character(df$adress)
df$electricity_costs_usd <- as.numeric(df$electricity_costs_usd)
df$borough <- as.factor(df$borough)
df$area_name <- as.factor(df$area_name)
```

```{r}
summary(df)
```

Time for some feature engineering of the adress column name, from that i can get building name and keep the adress.

```{r}
df$building_name <- NA

library(stringr)

df$building_name <- str_extract(df$adress, "\\[([^]]+)\\].*") #extracting clammers and whats inside them

df$building_name <- str_replace(df$building_name, "\\[([^]]+)\\].*", "\\1") #Removing clammers

df$building_name[df$building_name =='110 Chester St'] <- 'Dr. Rose B. English' #Renaming the building to the right name from my own manual research.

df$adress <- str_extract(df$adress, ".*(?=\\[)") #removes everything efter the first clammer and the clammer

df$adress <- trimws(df$adress) #Trimming leading and ending whitespaces
df$building_name <- trimws(df$building_name) #Trimming leading and ending whitespaces
df$area_name <- trimws(df$area_name) #Trimming leading and ending whitespaces

df$building_name <- gsub("?", "", df$building_name, fixed = TRUE) #Removing a trailing "?" that was after some building names

df$adress[df$adress =='111 Bristol St [110 Chester St]'] <- '111 Bristol St' #Fixing this manually
```

Now It's time to rename df$dates to df$date

```{r}
df$date <- df$dates
df$dates <- NULL
```

Checking if there are any duplicates

```{r}
n_occur <- data.frame(table(df$building_name))

n_occur[n_occur$Freq > 46,]
```
Giving both the names an unique identifier

```{r}
unique(df[df$building_name == 'Roberto Clemente School', c('building_name', 'area_name')])
```
```{r}
df$building_name[df$building_name == 'Roberto Clemente School' & df$area_name == 'Highbridge'] <- 'Roberto Clemente School	(Highbridge)'

df$building_name[df$building_name == 'Roberto Clemente School' & df$area_name == 'Kingsbridge Heights'] <- 'Roberto Clemente School	(Kingsbridge Heights)'
```

```{r}
remove(n_occur) #dont need this anymore
```


```{r}
summary(df)
```
It looks good! Time to export the cleaned data to a csv file, for further manipulation in excel!

```{r}
write.csv(df, file = 'cleaned_data.csv', row.names = F)
```

